---
title: "Temp-Nut Plots"
author: "Whitney Beck"
date: "February 12, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(car)
library(lsmeans)

Data=read.csv("raw-data/Beck Data/Beck_Longform.csv")

```

##1. ANCOVAs testing the effect of temperature on nutrient treatment (raw chlorophyll a) and nutrient LRRs
##Lsmeans tests for significant differences among treatments, averaging over temperature
##Lstrends tests for differences among the relationships between treatments and temperature 


### a. Temperature and raw Chl-a

```{r}

#Chla vs. Temp

ggplot(Data, aes(x=Temp, y=log(Chla_Mean+.01), color=Trt)) +
  geom_pointrange(aes(ymin=log(Chla_Mean+.01)-log(Chla_SE+.01), ymax=log(Chla_Mean+.01)+ log(Chla_SE+.01), shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(Log~Chlorophyll~italic(a)~mu*g*~cm^-2))+
  xlab(bquote(Temperature))+ 
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~Trt*Temp, data=Data)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over temperature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to temperature)

comp=lstrends(Model, ~ Trt, var = "Temp")
cld(comp)

```

### b.) Temperature and Nutrient LRRs

```{r cars}

#LRRs vs. Temp

ggplot(Data, aes(x=Temp, y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(Temperature))+ 
  ylim(-5,5) +
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*Temp, data=Data)
Anova(Model, type="III")

#Doesn't quite meet assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over temperature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to temperature)

comp=lstrends(Model, ~ Trt, var = "Temp")
cld(comp)



```

### c.) NO3_ug_L and raw Chl-a

```{r }

#Chla vs. NO3_ug_L

ggplot(Data, aes(x=log(NO3_ug_L+.01), y=log(Chla_Mean+.01), color=Trt)) +
  geom_pointrange(aes(ymin=log(Chla_Mean+.01)-log(Chla_SE+.01), ymax=log(Chla_Mean+.01)+ log(Chla_SE+.01), shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(Log~Chlorophyll~italic(a)~mu*g*~cm^-2))+
  xlab(bquote(Log~NO3_ug_L))+ 
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. NO3_ug_L ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~Trt*log(NO3_ug_L+.01), data=Data)
Anova(Model, type="III")

#Log meets assumptions pretty well

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over NO3_ug_Lerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to NO3_ug_Lerature)

comp=lstrends(Model, ~ Trt, var = "NO3_ug_L")
cld(comp)

```

### d.) NO3_ug_L and Nutrient LRRs

```{r }

#LRRs vs. NO3_ug_L

ggplot(Data, aes(x=log(NO3_ug_L+.01), y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(log(NO3_ug_L+.01)))+ 
  ylim(-5,5) +
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. NO3_ug_L ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*NO3_ug_L, data=Data)
Anova(Model, type="III")

#Doesn't really meet assumptions- outliers? 

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over NO3_ug_Lerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to NO3_ug_Lerature)

comp=lstrends(Model, ~ Trt, var = "NO3_ug_L")
cld(comp)

```

### e.) SRP_ug_L and raw Chl-a

```{r }

#Chla vs. SRP_ug_L

ggplot(Data, aes(x=log(SRP_ug_L+.01), y=log(Chla_Mean+.01), color=Trt)) +
  geom_pointrange(aes(ymin=log(Chla_Mean+.01)-log(Chla_SE+.01), ymax=log(Chla_Mean+.01)+ log(Chla_SE+.01), shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(Log~Chlorophyll~italic(a)~mu*g*~cm^-2))+
  xlab(bquote(Log~SRP_ug_L))+
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. SRP_ug_L ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~Trt*log(SRP_ug_L+.01), data=Data)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over SRP_ug_Lerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to SRP_ug_Lerature)

comp=lstrends(Model, ~ Trt, var = "SRP_ug_L")
cld(comp)

```

### f.) SRP_ug_L and Nutrient LRRs

```{r }

#LRRs vs. SRP_ug_L

ggplot(Data, aes(x=log(SRP_ug_L+.01), y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(log(SRP_ug_L+.01)))+ 
  ylim(-5,5)+
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. SRP_ug_L ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*log(SRP_ug_L+.01), data=Data)
Anova(Model, type="III")

#Doesn't really meet assumptions- outliers? 

par(mfrow=c(2,2))
plot(Model)

#Lsmeans (pairwise trt differences, averaging over SRP_ug_Lerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to SRP_ug_Lerature)

comp=lstrends(Model, ~ Trt, var = "SRP_ug_L")
cld(comp)

```

### g.) Canopy_Quant and raw Chl-a

```{r }

#Chla vs. Canopy_Quant

ggplot(Data, aes(x=Canopy_Quant, y=log(Chla_Mean+.01), color=Trt)) +
  geom_pointrange(aes(ymin=log(Chla_Mean+.01)-log(Chla_SE+.01), ymax=log(Chla_Mean+.01)+ log(Chla_SE+.01), shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(Log~Chlorophyll~italic(a)~mu*g*~cm^-2))+
  xlab(bquote(Log~Canopy_Quant))+ 
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. Canopy_Quant ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~Trt*Canopy_Quant, data=Data)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over Canopy_Quanterature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to Canopy_Quanterature)

comp=lstrends(Model, ~ Trt, var = "Canopy_Quant")
cld(comp)

```

### h.) Canopy_Quant and Nutrient LRRs

```{r }

#LRRs vs. Canopy_Quant

ggplot(Data, aes(x=Canopy_Quant, y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(log(Canopy_Quant+.01)))+ 
  ylim(-5,5)+
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. Canopy_Quant ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*Canopy_Quant, data=Data)
Anova(Model, type="III")

#Doesn't really meet assumptions- a few outliers

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over Canopy_Quanterature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to Canopy_Quanterature)

comp=lstrends(Model, ~ Trt, var = "Canopy_Quant")
cld(comp)

```

### i.) NO3_SRP_Ratio and raw Chl-a

```{r }

#Chla vs. NO3_SRP_Ratio

ggplot(Data, aes(x=log(NO3_SRP_Ratio+.01), y=log(Chla_Mean+.01), color=Trt)) +
  geom_pointrange(aes(ymin=log(Chla_Mean+.01)-log(Chla_SE+.01), ymax=log(Chla_Mean+.01)+ log(Chla_SE+.01), shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(Log~Chlorophyll~italic(a)~mu*g*~cm^-2))+
  xlab(bquote(Log~NO3_SRP_Ratio))+ 
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. NO3_SRP_Ratio ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~Trt*log(NO3_SRP_Ratio+.01), data=Data)
Anova(Model, type="III")

#Log generally meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over NO3_SRP_Ratioerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to NO3_SRP_Ratioerature)

comp=lstrends(Model, ~ Trt, var = "NO3_SRP_Ratio")
cld(comp)

```

### j.) NO3_SRP_Ratio and Nutrient LRRs

```{r }

#LRRs vs. NO3_SRP_Ratio

ggplot(Data, aes(x=log(NO3_SRP_Ratio+.01), y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(log(NO3_SRP_Ratio+.01)))+ 
  ylim(-5,5)+
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. NO3_SRP_Ratio ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*log(NO3_SRP_Ratio+.01), data=Data)
Anova(Model, type="III")

#Doesn't quite meet assumptions- outliers? 

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over NO3_SRP_Ratioerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to NO3_SRP_Ratioerature)

comp=lstrends(Model, ~ Trt, var = "NO3_SRP_Ratio")
cld(comp)

```


### k.) NO3_ug_L and Nutrient LRRs

```{r }

#LRRs vs. NO3_ug_L

ggplot(Data, aes(x=log(NO3_ug_L+.01), y=LRR_Mean, color=Trt)) +
  geom_pointrange(aes(ymin=LRR_Mean-LRR_Var, ymax=LRR_Mean + LRR_Var, shape=Trt, color=Trt))+
  geom_smooth(method=lm,   
                se=FALSE)+
    #scale_colour_hue(l=50) + 
  theme(axis.text.x = element_text(colour="black"), panel.grid.minor = element_blank(), panel.grid.major = element_blank(), axis.text.y=element_text(colour="black"), panel.background=element_blank(), plot.title=element_text(hjust=-.22))+
  ylab(bquote(LRR))+
  xlab(bquote(log(NO3_ug_L+.01)))+ 
  ylim(-5,5)+
  scale_colour_manual(values=c("#377eb8", "#4daf4a", "#984ea3", "#ff7f00"))

#Chla vs. NO3_ug_L ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~Trt*log(NO3_ug_L+.01), data=Data)
Anova(Model, type="III")

#Meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

#Lsmeans (pairwise trt differences, averaging over NO3_ug_Lerature)

lsmeans(Model, pairwise~Trt)

#Lstrends (pairwise trt differences in the way chl-a responds to NO3_ug_Lerature)

comp=lstrends(Model, ~ Trt, var = "NO3_ug_L")
cld(comp)

```

## 2.) Three-way interaction models examine the effects of in-stream NO3, SRP, and temp on chlorophyll a (broken out by treatment) and LRRs (broken out by treatment)

### a.) Temp-Nut Interactions and raw Chl-a

```{r }

##Controls

Ctrl=subset(Data, Trt=="Ctrl")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=Ctrl)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)


##Nitrogens

Nitrogen=subset(Data, Trt=="N")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=Nitrogen)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)


##Phosphorus

Phosphorus=subset(Data, Trt=="P")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=Phosphorus)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)


##Nitrogen + Phosphorus

NP=subset(Data, Trt=="NP")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(log(Chla_Mean+.01)~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=NP)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)
```

### b.) Temp-Nut interactions on LRRs

```{r }

##Nitrogens

Nitrogen=subset(Data, Trt=="N")

#LRR vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=Nitrogen)
Anova(Model, type="III")

#Log mostly meets assumptions

#par(mfrow=c(2,2))
#plot(Model)


##Phosphorus

Phosphorus=subset(Data, Trt=="P")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=Phosphorus)
Anova(Model, type="III")

#Log mostly meets assumptions

#par(mfrow=c(2,2))
#plot(Model)


##Nitrogen + Phosphorus

NP=subset(Data, Trt=="NP")

#Chla vs. NO3_ug_L, SRP, Temp ANCOVAs

options(contrasts=c("contr.sum", "contr.poly"))
Model=aov(LRR_Mean~log(NO3_ug_L+.01)*log(SRP_ug_L)*Temp, data=NP)
Anova(Model, type="III")

#Log meets assumptions

#par(mfrow=c(2,2))
#plot(Model)

```